\section{KC3DCONN: The Connector Toolbox}
The module \textbf{kc3dconn} contains parametric models
of connectors.  We have already seen an example of such
a model in Section~\ref{sec:quick-example}.

\subsection{kc3dconn:Genhdr}
The class \textbf{Genhdr} represents a generic square-pitch vertical header.
The header may have square or circular pins, may be male or female, may
have a plain or beveled casing, and may have a shoulder on the casing.
Section~\ref{sec:quick-example} presented this class and demonstrated its
use although no details were discussed. Internally the class makes use of
the classes \textbf{VRMLMat}, \textbf{Hole}, \textbf{Hdrbase}, \textbf{Pin},
\textbf{Funnel}, and \textbf{Circle} to generate all the features of the
header.

The parameters controlling the final appearance of the header have been
divided into a number of groups: Color parameters, Case parameters, and
Pin parameters.

The colors can be selected by invoking \textbf{setColors(case, pin\_out, pin\_in, shroud)}
where each argument is of type VRMLMat, case is the color of the case, pin\_out is the outer
appearance of the pin, and pin\_in is the inner appearance of the pin, and shroud is the appearance of
the recess or protrusion around a pin; pin\_out and shroud are applicable only
to female headers.

The case parameters are set via \textbf{setCase(columns, rows, col\_pitch, row\_pitch, height, shoulder, hassh, bevel)}
where height is the total case height, shoulder is the height of the case shoulder, and bevel is the
case bevel between columns. The parameter hassh controls whether a shoulder is rendered (True) or if the shoulder
parameter simply specifies a gap between teh PCB and the case (False).

The pin parameters are set via
\textbf{setPins(squarebot, squaretop, male, pbev, fbev, depth, length, pd0, pdy, pd1, pd2, pd3, ftc, taper, ts, sides, funneldepth)}:

\begin{itemize}
\item \textbf{squarebot}: set True if the pin is square or rectangular
\item \textbf{squaretop}: set True if the receptacle is square (applicable only to female headers)
\item \textbf{male}: set True if the header is male
\item \textbf{pbev}: pin bevel; applicable only if squarebot is True
\item \textbf{fbev}: funnel bevel; applicable only if male is False and squaretop is True
\item \textbf{depth}: pin length below the top of the PCB
\item \textbf{length}: total pin length (depth + total connector height above PCB)
\item \textbf{pd0}: X dimension of pins
\item \textbf{pdy}: Y dimension of pins; applicable only if squarebot is True
\item \textbf{pd1}: dimension of the shroud around the receptacle; applicable only when male is False
\item \textbf{pd2}: inner dimension of the receptacle; applicable only when male is False
\item \textbf{pd3}: diameter of the pin where it enters the bottom of the casing; applicable only when
    male and squarebot are False
\item \textbf{ftc}: when male and squaretop are False, ftc controls the bevel at the receptacle's entrance;
    valid values are $\ge1$.
\item \textbf{taper}: pin taper length (set to $\le0$ for none
\item \textbf{ts}: taper coefficient; controls the size of the tapered end
\item \textbf{sides}: number of vertices to use for circular pins or receptacles
\item \textbf{funneldepth}: for a square receptacle this is the depth of the funnel's flute and must be $\ge0$;
    for a circular receptacle this is the depth of the recess ($>0$), the height of the protrusion ($<0$),
    or 0 for a receptacle which is flush with the housing.
\end{itemize}

Genhdr has been used to create models for Samtec headers in series SS, HSS, ESS, SD, ESD, SL, and SLD
and can be used to generate models for headers from many other manufacturers.
